Abstractions for Multi-Sorted Substitutions

### 2024-03-13 paper/paper.tex
Update.

  In this section, we extend our framework with an abstraction that
  allows the composition of arbitrary maps.
  The core property of composition, which we call
  \ACode{\AField{⋯-fusion}}, states that applying two maps \ACode{ϕ₁}
  and \ACode{ϕ₂} in sequence to a term \ACode{t}, is the same as
  applying their composition \ACode{(ϕ₁~\AField{·ₖ}~ϕ₂)} to \ACode{t},
  i.e. \ACode{(t \AField{⋯} ϕ₁) \AField{⋯} ϕ₂ \ACon{≡} t \AField{⋯} (ϕ₁ \AField{·ₖ} ϕ₂)}.
  % $$
  % \ACode{(t \AField{⋯} ϕ₁) \AField{⋯} ϕ₂ \ACon{≡} t \AField{⋯} (ϕ₁ \AField{·ₖ} ϕ₂)}
  % $$
  This property equips our framework with the power to
  reason about the application of multiple maps by reasoning about
  the application of a single map.
  As such it forms the basis for all lemmas involving multiple maps,
  e.g.\ that applying a weakening and then a singleton
  substitution cancel each other out.

@@

  As we defined substitution in terms of renamings, we need to
  consider all four compositions between renamings and substitutions.
  While the composition operations can be defined independently of each other,
  the \ACode{\AField{⋯-fusion}} lemma for two substitutions, depends on the
  \ACode{\AField{⋯-fusion}} lemmas between a renaming and a substitution, which in
  turn depend on the \ACode{\AField{⋯-fusion}} lemma between two renamings.

@@

  % \ACode{\ACon{CKit}} and \ACode{\ADef{\_·ₖ\_}} are analogous to
  % \ACode{\ACon{Kit}} and \ACode{\ADef{\_→ₖ\_}}, whereas
  % \ACode{\ACon{CTraversal}} and \ACode{\AField{⋯-fusion}} are analogous to
  % \ACode{\ACon{CTraversal}} and \ACode{\AField{\_⋯\_}}.

@@

  In this subsection, we define the \ACode{\ACon{CKit}}
  abstraction, which provides the \ACode{\AField{\_\&/⋯\_}} operation
  that is necessary for defining composition, and the
  \ACode{\ACon{CTraversal}}, which provides the
  \ACode{\AField{⋯-fusion}} lemma for general maps.

  Similarly as it was the case for \ACode{\ACon{Kit}} and \ACode{\ACon{Traversal}},
  the lemma of the \ACode{\ACon{CTraversal}} needs to be proved for
  an abstract \ACode{\ACon{CKit}} and in return defines the
  four concrete \ACode{\ACon{CKit}} instances for us. This allows us to break
  the dependencies between the different \ACode{\AField{⋯-fusion}} lemmas,
  similar to how \ACode{\ACon{Kit}} and \ACode{\ACon{Traversal}} allowed
  us to break the dependency between traversals with renamings and substitutions.

@@

  We then define a structure,
  which provides the \ACode{\AField{⋯-fusion}} lemma for arbitrary
  \ACode{\ACon{CKit}}s:
  \LibCode\KCTraversal

@@

    % A map \ACode{ϕ} followed by a weakening is equivalent to a
    % weakening followed by \ACode{ϕ} that has been lifted over the weakened
    % variable:
    % \LibCode\KCommLiftWeaken
    % % \LibCode\KCommLiftWeakenProof
    % By applying \ACode{\AField{⋯-fusion}} to this lemma, we get the same
    % statement for multiple map applications:
    % \LibCode\KCommLiftWeakenTraverse
    % % \LibCode\KCommLiftWeakenTraverseProof
  \item 
    A weakening followed by a singleton substitution cancel each other out:
    \LibCode\KWeakenCancelsSingleTraverse

    % A weakening followed by a singleton substitution cancel each other out:
    % \LibCode\KWeakenCancelsSingle
    % % \LibCode\KWeakenCancelsSingleProof
    % By applying \ACode{\AField{⋯-fusion}} and \ACode{\AField{⋯-id}} to
    % this lemma, we get the same statement for multiple map applications:
    % \LibCode\KWeakenCancelsSingleTraverse
    % % \LibCode\KWeakenCancelsSingleTraverseProof
  \item 
    A singleton map can be swapped with any map \ACode{ϕ}:
    \LibCode\KDistLiftSingleTraverse
    % A singleton map can be swapped with any map \ACode{ϕ}:
    % \LibCode\KDistLiftSingle
    % % \LibCode\KDistLiftSingleProof
    % By applying \ACode{\AField{⋯-fusion}} to this lemma, we get the
    % same statement for multiple map applications:
    % \LibCode\KDistLiftSingleTraverse
    % % \LibCode\KDistLiftSingleTraverseProof
  \item 
    We define \ACode{\ACon{CKit}} instances for each of the four
    combinations between renamings and substitutions:
    \LibCode\KCKitInstances
    Interestingly, the \ACode{\ADef{Cₛ}} instance has another
    \ACode{\ADef{CKit}} parameter.
    This allows us to first use \ACode{\ADef{Cᵣ}} as an argument,
    which yields the \ACode{\ADef{CKit}} between substitutions
    and renamings, and then use this resulting \ACode{\ADef{CKit}}
    as an argument for the case with two substitutions.
    As shown below, Agda's instance resolution is up for the task:

>>>>

  In this section, we extend our framework with an abstraction that
  allows the composition of arbitrary maps.
  The core property of composition is the
  \ACode{\AField{fusion}} lemma, which states that applying two maps \ACode{ϕ₁}
  and \ACode{ϕ₂} in sequence to a term \ACode{t}, is the same as
  applying their composition \ACode{(ϕ₁~\AField{·ₖ}~ϕ₂)} to \ACode{t},
  i.e. \ACode{(t \AField{⋯} ϕ₁) \AField{⋯} ϕ₂ \ACon{≡} t \AField{⋯} (ϕ₁ \AField{·ₖ} ϕ₂)}.
  % $$
  % \ACode{(t \AField{⋯} ϕ₁) \AField{⋯} ϕ₂ \ACon{≡} t \AField{⋯} (ϕ₁ \AField{·ₖ} ϕ₂)}
  % $$
  This property gives our framework the ability to
  reason about the application of multiple maps by reasoning about
  the application of a single map.
  As such it forms the basis for all lemmas involving multiple maps,
  e.g.\ that applying a weakening and then a singleton
  substitution cancel each other out.

@@

  As we defined substitution in terms of renamings, we need to
  consider all four compositions between renamings and substitutions.
  While the composition operations can be defined independently of each other,
  the \ACode{\AField{fusion}} lemma for two substitutions, depends on the
  \ACode{\AField{fusion}} lemmas between a renaming and a substitution, which in
  turn depend on the \ACode{\AField{fusion}} lemma between two renamings.

@@

  % \ACode{\ACon{CKit}} and \ACode{\ADef{\_·ₖ\_}} are analogous to
  % \ACode{\ACon{Kit}} and \ACode{\ADef{\_→ₖ\_}}, whereas
  % \ACode{\ACon{CTraversal}} and \ACode{\AField{fusion}} are analogous to
  % \ACode{\ACon{CTraversal}} and \ACode{\AField{\_⋯\_}}.

@@

  In this subsection, we define the \ACode{\ACon{CKit}}
  abstraction, which provides the \ACode{\AField{\_\&/⋯\_}} operation
  that is necessary for defining composition, and the
  \ACode{\ACon{CTraversal}}, which provides the
  \ACode{\AField{fusion}} lemma for general maps.

  Similarly as it was the case for \ACode{\ACon{Kit}} and \ACode{\ACon{Traversal}},
  the lemma of the \ACode{\ACon{CTraversal}} needs to be proved for
  an abstract \ACode{\ACon{CKit}} and in return defines the
  four concrete \ACode{\ACon{CKit}} instances for us. This allows us to break
  the dependencies between the different \ACode{\AField{fusion}} lemmas,
  similar to how \ACode{\ACon{Kit}} and \ACode{\ACon{Traversal}} allowed
  us to break the dependency between traversals with renamings and substitutions.

@@

  We then define a structure,
  which provides the \ACode{\AField{fusion}} lemma for arbitrary
  \ACode{\ACon{CKit}}s:
  \LibCode\KCTraversal

@@

    % A map \ACode{ϕ} followed by a weakening is equivalent to a
    % weakening followed by \ACode{ϕ} that has been lifted over the weakened
    % variable:
    % \LibCode\KCommLiftWeaken
    % % \LibCode\KCommLiftWeakenProof
    % By applying \ACode{\AField{fusion}} to this lemma, we get the same
    % statement for multiple map applications:
    % \LibCode\KCommLiftWeakenTraverse
    % % \LibCode\KCommLiftWeakenTraverseProof
  \item 
    A weakening followed by a singleton substitution cancel each other out:
    \LibCode\KWeakenCancelsSingleTraverse

    % A weakening followed by a singleton substitution cancel each other out:
    % \LibCode\KWeakenCancelsSingle
    % % \LibCode\KWeakenCancelsSingleProof
    % By applying \ACode{\AField{fusion}} and \ACode{\AField{⋯-id}} to
    % this lemma, we get the same statement for multiple map applications:
    % \LibCode\KWeakenCancelsSingleTraverse
    % % \LibCode\KWeakenCancelsSingleTraverseProof
  \item 
    A singleton map can be swapped with any map \ACode{ϕ}:
    \LibCode\KDistLiftSingleTraverse
    % A singleton map can be swapped with any map \ACode{ϕ}:
    % \LibCode\KDistLiftSingle
    % % \LibCode\KDistLiftSingleProof
    % By applying \ACode{\AField{fusion}} to this lemma, we get the
    % same statement for multiple map applications:
    % \LibCode\KDistLiftSingleTraverse
    % % \LibCode\KDistLiftSingleTraverseProof
  \item 
    We define \ACode{\ACon{CKit}} instances for each of the four
    combinations between renamings and substitutions:
    \LibCode\KCKitInstances
    Interestingly, the \ACode{\ADef{Cₛ}} instance has another
    \ACode{\ADef{CKit}} parameter.
    This allows us to first use \ACode{\ADef{Cᵣ}} as an argument,
    which yields the \ACode{\ADef{CKit}} between substitutions
    and renamings, and then use this resulting \ACode{\ADef{CKit}}
    as an argument for the case with two substitutions.
    As shown below, Agda's instance resolution is up for the task:
